---
title: "01 Prep"
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
bibliography: "../chen-2020-ST.bib"
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "results", "01-prep.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [chen-2020-ST](https://github.com/eturkes/chen-2020-ST).*
*Please email for access.*

In this document we prepare the gene count matrix for downstream analysis.
The data here is derived from @`r unlist(strsplit(getwd(), "/"))[4]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[4]``.

```{r}
# Some standard boilerplate.
# --------------------------
#    This file is part of chen-2020-ST.
#    Copyright (C) 2019-2020  Emir Turkes, UK DRI at UCL
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# This section should be checked per document.
# --------------------------------------------
packages <- c("data.table", "DropletUtils", "Seurat")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))

setDTthreads(0)
analysis_no <- 1
# --------------------------------------------

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------
data_name <- unlist(strsplit(getwd(), "/"))[4] # Name of dataset.
assets_dir <- file.path("..", "assets") # Backed up data.

cache_dir <- file.path("..", "tmp", "cache")
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 7)
# ----------------------------------------------------------
# --------------------------
```

# Cleaning

```{r}
counts <- t(fread(file.path(assets_dir, "data", "raw_counts.txt")))
colnames(counts) <- counts[1, ]
counts <- counts[-1, ]
counts <- counts[complete.cases(counts), ]
counts <- Matrix::Matrix(counts, sparse = TRUE)

meta <- read.delim(file.path(assets_dir, "data", "spot_metadata.tsv"))
meta <- meta[meta$Spot %in% colnames(counts), ]
counts <- counts[ , colnames(counts) %in% meta$Spot]
meta <- meta[match(colnames(counts), meta$Spot), ]

gc()
```

# B02_D1

```{r}
sub_name <- "B02_D1"

keep <- meta[meta$SampleID == sub_name, 1]
B02_D1_counts <- counts[ , colnames(counts) %in% keep]
B02_D1_meta <- meta[meta$Spot %in% colnames(B02_D1_counts), ]

tissue_positions_list <- B02_D1_meta[ , c(1, 12, 15:18)]
tissue_positions_list$Region_equalsize <- ifelse(
  tissue_positions_list$Region_equalsize == "", 0, 1
)

if (
  !dir.exists(file.path(assets_dir, "data", sub_name, "raw_feature_bc_matrix"))
) {
  write10xCounts(
    file.path(assets_dir, "data", sub_name, "raw_feature_bc_matrix"),
    B02_D1_counts,
    type = "sparse",
    version = "3"
  )
  write10xCounts(
    file.path(
      assets_dir, "data", sub_name,
      "raw_feature_bc_matrix", "filtered_feature_bc_matrix.h5"
    ),
    B02_D1,
    type = "HDF5",
    version = "3"
  )
  write.table(
    tissue_positions_list,
    file.path(
      assets_dir, "data", sub_name,
      "raw_feature_bc_matrix", "spatial", "tissue_positions_list.csv"
    ),
    quote = FALSE,
    row.names = FALSE,
    col.names = FALSE,
    sep = ","
  )
}

seurat <- Load10X_Spatial(
  file.path(assets_dir, "data", sub_name, "raw_feature_bc_matrix")
)
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
